---
title: "Welcome to the workshop"
format: qmdrslides-revealjs
editor_options: 
  chunk_output_type: inline
---

```{r}
#| echo: false
library(countdown)
```

##  {data-menu-item="Workshop Goals"}

\
\

### Goals for this session {style="font-size: 2.5em; text-align: center"}

::: {.incremental style="font-size: 1.5em"}
1.  Get familiar with R-Studio

2.  Set expectations for today

3.  Get excited!
:::

##  {data-menu-title="Website Link" style="text-align: center"}

\
\
\

::: r-fit-text
Workshop materials are at:

<https://ecotopic.github.io/Rworkshop/>
:::

## Practicalities

::: r-fit-text
WiFi:

Network: KTB Free Wifi (no password needed)



Bathrooms are out the lobby, opposite this venue.
:::

```{r}
#| echo: false

#Network AHRI Password: `@hR1W1F1!17`
#Network CAPRISA-Corp Password: `corp@caprisa17`
```


````{=html}
<!--
## Introductions {.your-turn}

```{r}
#| echo: false
countdown::countdown(3)
```

Take \~3 minutes to introduce yourself to your neighbors.


Please share ...

1.  Your name
2.  Where you're from and where you work
3.  Your current go-to method for analyzing data
-->
````

## Your Instructors {.our-turn}

Who are we?

-   Johan van der Molen (CAPRISA)
-   Aaron Abai (Ragon Institute of MGH, MIT and Harvard)
-   Dr. Afrah Khairallah (AHRI)
-   Bertha Baye (AHRI)
-   Asavela Kama (CAPRISA)

## Let's make this workshop work for all

::: {.incremental style="font-size: 1.2em"}
-   You belong here. This workshop is intended for a wide-audience with a focus on beginners. If you feel out of place - it's our problem, not yours!

-   Stay committed. This workshop is intended to show you some data-analysis tools and leave you with skills you can really use. With practice you will hone these skills. Commit to stay engaged for best results, for you and those around you!


:::

## The premise of the workshop

We will be using a fictional but true-to-life "medium dimensional" dataset that we'll use for all instruction

-   This is a fictional Ebola outbreak, expanded by the handbook team from the ebola_sim practice dataset in the outbreaks package.
-   From this website: [The Epidemiologist R Handbook](https://www.epirhandbook.com/en/index.html)

Why this format

- Hands on.
- This is the first time running this new workshop - please give lots of feedback on how to improve it for next time.
- However, many of you in the room know a lot more immunology/epidemiology - and even if you know none you can learn R and generalize it to your research.

## The content of this workshop

::: {style="font-size: 1.2em"}
This is probably too much material to get through today!\
\
As instructors we're going to be trying to teach at the right pace to keep everyone learning.\
\
The materials will stay on the website forever for you to work through at your own pace.
:::
